Manuscript Figures

### 2018-06-14 Manuscript Figures.Rmd
Includes basic text output for several counts used in manuscript

```{r importSynMapData, include=FALSE}
source("./R/do.R")

@@

@@

```{r chromosome_plot_abundance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=8, fig.height=6}
subgenome.positions <-
  subgenome %>%
  left_join(B73.v4.gene.positions, by=c("gene2"="geneID")) %>%
  select(Chromosome, Pos_start, Pos_end, subgenome) %>%
  rename(Subgenome=subgenome) %>%
  subset(Subgenome %in% c("sub1", "sub2"))

dominant_homeolog_abun <- expressedProteinsPairs
dominant_homeolog_abun$Subgenome <- "no dominance"
dominant_homeolog_abun$Subgenome[dominant_homeolog_abun$dNSAF_mean1 / dominant_homeolog_abun$dNSAF_mean2 > 2] <- "Dominant Abundance"
dominant_homeolog_abun$Subgenome[dominant_homeolog_abun$dNSAF_mean2 / dominant_homeolog_abun$dNSAF_mean1 > 2] <- "Dominant Abundance"
dominant_homeolog_abun$Maize1[dominant_homeolog_abun$dNSAF_mean2 / dominant_homeolog_abun$dNSAF_mean1 > 2] <- dominant_homeolog_abun$Maize2[dominant_homeolog_abun$dNSAF_mean2 / dominant_homeolog_abun$dNSAF_mean1 > 2]
df0 <- dominant_homeolog_abun[dominant_homeolog_abun$Subgenome %in% c("no dominance"),]
df0$Maize1 <- df0$Maize2
dominant_homeolog_abun <- bind_rows(dominant_homeolog_abun, df0)
dominant_homeolog_abun <-
  dominant_homeolog_abun %>%
  select(Maize1, Subgenome) %>%
  left_join(B73.v4.gene.positions, by=c("Maize1"="geneID")) %>%
  select(Chromosome, Pos_start, Pos_end, Subgenome)

df1 <- B73.v4.centromere.positions
df1$x <- -.45
df1$xend <- .45
df2 <- subgenome.positions
df2$x <- -.45
df2$xend <- -.05
df3 <- dominant_homeolog_abun
df3$x <- .05
df3$xend <- .45
segment_data <- bind_rows(df2,df3)
segment_data <- segment_data[!is.na(segment_data$Pos_start),]
segment_data <-
  segment_data %>%
  ungroup()
segment_data$Subgenome[segment_data$Subgenome %in% c("sub1")] <- "Maize1"
segment_data$Subgenome[segment_data$Subgenome %in% c("sub2")] <- "Maize2"
# segment_data$Subgenome[segment_data$Subgenome %in% c("dom")] <- "Dominant Homeolog"
segment_data <- segment_data %>% subset(Subgenome %in% c("Maize1", "Maize2", "Dominant Abundance"))
centromere_data <- df1

## Make the plot
chromosome_plot_abundance <-
  ggplot(data=B73.v4.chr.size) +
  geom_bar(aes(B73.v4.chr.size$chromosome, B73.v4.chr.size$size), stat="identity", fill="grey70") +
  # geom_segment(data=segment_data, aes(x=Chromosome+x, xend=Chromosome+xend, y=Pos_start, yend=Pos_end, colour=Subgenome), size=.001) +
  geom_rect(data=segment_data, inherit.aes = F, aes(xmin=Chromosome+x, xmax=Chromosome+xend, ymin=Pos_start, ymax=Pos_end, color=Subgenome, fill=Subgenome)) +
  geom_rect(data=centromere_data, inherit.aes = F, aes(xmin=Chromosome+x, xmax=Chromosome+xend, ymin=Pos_start, ymax=Pos_end)) +
  scale_fill_manual(values=c("green", "red", "blue", "black")) +
  scale_color_manual(values=c("green", "red", "blue", "black")) +
  scale_x_discrete(limits=c("Chr1","Chr2","Chr3","Chr4","Chr5","Chr6","Chr7","Chr8","Chr9","Chr10")) +
  scale_y_continuous(labels=format_si())
chromosome_plot_abundance  +
  labs(
    # title="Subgenome Locations by Chromosome",
    x="Chromosome",
    y="Position on Chromosome"
  )
```

@@

```{r extra6, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.height=4, fig.align='center'}
## Not really a plot yet... just sorting types
source("./R/analysis/functionalDivergence_func.R")
"gene"
gene.pair.types <- getTypeSort(homeologs.pairs, maize.walley.v4mapped.expression, 2, .95)
summary(as.factor(gene.pair.types$type))
"protein"
protein.pair.types <- getTypeSortProteins(homeologs.pairs, maize.walley.abundance.v4, 2, .95)
summary(as.factor(protein.pair.types$type))
"agree"
summary(as.factor(intersect(gene.pair.types, protein.pair.types)$type))

# intersect(gene.pair.types[!gene.pair.types$type %in% c("ambiguous"),], protein.pair.types[!protein.pair.types$type %in% c("ambiguous"),])
# #dead expression and still dead abundance (17/297)
# protein.pair.types[protein.pair.types$Maize1 %in% gene.pair.types$Maize1[gene.pair.types$type %in% c("1dead","2dead")],] %>% subset(type %in% c("1dead","2dead"))
```

@@

```{r print, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
# # ggsave("Output/Figure 1a.png", plot = chromosome_plot_exp)
# png(filename = "Output/Figure 1a.png", type = "cairo", units = "in", width = 8, height = 8, pointsize = 12, res = 96)
# plot(chromosome_plot_exp)
# dev.off()
# 
# # ggsave("Output/Figure 1b.png", plot = chromosome_plot_abundance)
# png(filename = "Output/Figure 1b.png", type = "cairo", units = "in", width = 8, height = 8, pointsize = 12, res = 96)
# plot(chromosome_plot_abundance)
# dev.off()
# 
# png("Output/Figure 2a.png");
# grid.draw(GOTerms_by_subgenome_a)
# dev.off()
# png("Output/Figure 2b.png");
# grid.draw(GOTerms_by_subgenome_b)
# dev.off()
# 
# # ggsave("Output/Figure 3.png", plot = subgenome_isoforms_scatter)
# png(filename = "Output/Figure 3.png", type = "cairo", units = "in", width = 8, height = 8, pointsize = 12, res = 96)
# plot(subgenome_isoforms_scatter)
# dev.off()
# 
# # ggsave("Output/Figure 4.png", plot = subgenome_expression_scatter_rep)
# png(filename = "Output/Figure 4.png", type = "cairo", units = "in", width = 12, height = 4, pointsize = 12, res = 96)
# plot(subgenome_expression_scatter_rep)
# dev.off()
# 
# # ggsave("Output/Figure 5.png", plot = subgenome_protein_scatter_rep)
# png(filename = "Output/Figure 5.png", type = "cairo", units = "in", width = 12, height = 4, pointsize = 12, res = 96)
# plot(subgenome_protein_scatter_rep)
# dev.off()
# 
# # ggsave("Output/Figure 6.png", plot = subgenome_expression_protein_scatter_rep)
# png(filename = "Output/Figure 6.png", type = "cairo", units = "in", width = 12, height = 4, pointsize = 12, res = 96)
# plot(subgenome_expression_protein_scatter_rep)
# dev.off()
# 
# ggsave("Output/Figure S1.png", plot = subgenome_expression_scatter)
# ggsave("Output/Figure S2.png", plot = subgenome_protein_scatter)
# ggsave("Output/Figure S3.png", plot = subgenome_expression_protein_scatter)
# ggsave("Output/Figure S4.png", plot = subgenome_expression_protein_foldchange_scatter)
```

>>>>

```{r importSynMapData, include=FALSE}
source("./R/do.R")
source("./R/pathway_analysis.R")

@@

Some direct numbers for reporting in various parts of the manuscript.

Overlap in reactions and pathways annotated to retained pairs.

```{r numbers1, echo=FALSE, results='asis'}
knitr::kable(getPathwayReactionOverlap(homeologs.pairs) %>% select(Description, Value))
```

How often does the maize1 gene express higher than maize2?  Only count when both genes have expression.

```{r numbers2, echo=FALSE, results='asis'}
data <- 
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>% #only matching
  select(Maize1, Maize2) %>%
  inner_join(maize.walley.v4mapped.expression, by=c("Maize1"="geneID")) %>%
  rename(FPKM_avg1=FPKM_avg) %>%
  inner_join(maize.walley.v4mapped.expression, by=c("Maize2"="geneID", "sample"="sample")) %>%
  rename(FPKM_avg2=FPKM_avg) %>%
  select(Maize1, Maize2, sample, FPKM_avg1, FPKM_avg2) %>%
  count(sample, sign=sign(FPKM_avg1-FPKM_avg2)) %>%
  spread(key = sign, value = n) %>%
  mutate(percent_favor_m1=`1`/(`-1`+`1`)) %>%
  rename(m2_dom=`-1`, m1_dom=`1`) %>%
  select(sample, m1_dom, m2_dom, percent_favor_m1) %>%
  arrange(percent_favor_m1)
knitr::kable(data)
```

 * There are **`r nrow(homeologs.pairs %>% subset(Maize1 != "" & Maize2 != ""))`** retained duplicate pairs.
 * 
 <!-- * Summary of the KS values -->
 <!--     + **`r 1+2`** -->

@@

<!-- # ```{r chromosome_plot_abundance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=8, fig.height=6} -->
<!-- # subgenome.positions <- -->
<!-- #   subgenome %>% -->
<!-- #   left_join(B73.v4.gene.positions, by=c("gene2"="geneID")) %>% -->
<!-- #   select(Chromosome, Pos_start, Pos_end, subgenome) %>% -->
<!-- #   rename(Subgenome=subgenome) %>% -->
<!-- #   subset(Subgenome %in% c("sub1", "sub2")) -->
<!-- #  -->
<!-- # dominant_homeolog_abun <- expressedProteinsPairs -->
<!-- # dominant_homeolog_abun$Subgenome <- "no dominance" -->
<!-- # dominant_homeolog_abun$Subgenome[dominant_homeolog_abun$dNSAF_mean1 / dominant_homeolog_abun$dNSAF_mean2 > 2] <- "Dominant Abundance" -->
<!-- # dominant_homeolog_abun$Subgenome[dominant_homeolog_abun$dNSAF_mean2 / dominant_homeolog_abun$dNSAF_mean1 > 2] <- "Dominant Abundance" -->
<!-- # dominant_homeolog_abun$Maize1[dominant_homeolog_abun$dNSAF_mean2 / dominant_homeolog_abun$dNSAF_mean1 > 2] <- dominant_homeolog_abun$Maize2[dominant_homeolog_abun$dNSAF_mean2 / dominant_homeolog_abun$dNSAF_mean1 > 2] -->
<!-- # df0 <- dominant_homeolog_abun[dominant_homeolog_abun$Subgenome %in% c("no dominance"),] -->
<!-- # df0$Maize1 <- df0$Maize2 -->
<!-- # dominant_homeolog_abun <- bind_rows(dominant_homeolog_abun, df0) -->
<!-- # dominant_homeolog_abun <- -->
<!-- #   dominant_homeolog_abun %>% -->
<!-- #   select(Maize1, Subgenome) %>% -->
<!-- #   left_join(B73.v4.gene.positions, by=c("Maize1"="geneID")) %>% -->
<!-- #   select(Chromosome, Pos_start, Pos_end, Subgenome) -->
<!-- #  -->
<!-- # df1 <- B73.v4.centromere.positions -->
<!-- # df1$x <- -.45 -->
<!-- # df1$xend <- .45 -->
<!-- # df2 <- subgenome.positions -->
<!-- # df2$x <- -.45 -->
<!-- # df2$xend <- -.05 -->
<!-- # df3 <- dominant_homeolog_abun -->
<!-- # df3$x <- .05 -->
<!-- # df3$xend <- .45 -->
<!-- # segment_data <- bind_rows(df2,df3) -->
<!-- # segment_data <- segment_data[!is.na(segment_data$Pos_start),] -->
<!-- # segment_data <- -->
<!-- #   segment_data %>% -->
<!-- #   ungroup() -->
<!-- # segment_data$Subgenome[segment_data$Subgenome %in% c("sub1")] <- "Maize1" -->
<!-- # segment_data$Subgenome[segment_data$Subgenome %in% c("sub2")] <- "Maize2" -->
<!-- # # segment_data$Subgenome[segment_data$Subgenome %in% c("dom")] <- "Dominant Homeolog" -->
<!-- # segment_data <- segment_data %>% subset(Subgenome %in% c("Maize1", "Maize2", "Dominant Abundance")) -->
<!-- # centromere_data <- df1 -->
<!-- #  -->
<!-- # ## Make the plot -->
<!-- # chromosome_plot_abundance <- -->
<!-- #   ggplot(data=B73.v4.chr.size) + -->
<!-- #   geom_bar(aes(B73.v4.chr.size$chromosome, B73.v4.chr.size$size), stat="identity", fill="grey70") + -->
<!-- #   # geom_segment(data=segment_data, aes(x=Chromosome+x, xend=Chromosome+xend, y=Pos_start, yend=Pos_end, colour=Subgenome), size=.001) + -->
<!-- #   geom_rect(data=segment_data, inherit.aes = F, aes(xmin=Chromosome+x, xmax=Chromosome+xend, ymin=Pos_start, ymax=Pos_end, color=Subgenome, fill=Subgenome)) + -->
<!-- #   geom_rect(data=centromere_data, inherit.aes = F, aes(xmin=Chromosome+x, xmax=Chromosome+xend, ymin=Pos_start, ymax=Pos_end)) + -->
<!-- #   scale_fill_manual(values=c("green", "red", "blue", "black")) + -->
<!-- #   scale_color_manual(values=c("green", "red", "blue", "black")) + -->
<!-- #   scale_x_discrete(limits=c("Chr1","Chr2","Chr3","Chr4","Chr5","Chr6","Chr7","Chr8","Chr9","Chr10")) + -->
<!-- #   scale_y_continuous(labels=format_si()) -->
<!-- # chromosome_plot_abundance  + -->
<!-- #   labs( -->
<!-- #     # title="Subgenome Locations by Chromosome", -->
<!-- #     x="Chromosome", -->
<!-- #     y="Position on Chromosome" -->
<!-- #   ) -->
<!-- # ``` -->

@@

```{r extra6, echo=FALSE, results='asis', warning=FALSE, message=FALSE, fig.width=4, fig.height=4, fig.align='center'}
## Not really a plot yet... just sorting types
source("./R/analysis/functionalDivergence_func.R")
"gene"
gene.pair.types <- getTypeSort(homeologs.pairs, maize.walley.v4mapped.expression, 2, .95)
summary(as.factor(gene.pair.types$type))
"protein"
protein.pair.types <- getTypeSortProteins(homeologs.pairs, maize.walley.abundance.v4, 2, .95)
summary(as.factor(protein.pair.types$type))
"agree"
summary(as.factor(intersect(gene.pair.types, protein.pair.types)$type))

data <- 
  data.frame(type=c("gene","protein","agree"),
             values=bind_rows(summary(as.factor(gene.pair.types$type)), 
                              summary(as.factor(protein.pair.types$type)),
                              summary(as.factor(intersect(gene.pair.types, protein.pair.types)$type))
                              )
             ) %>% rename(one_dead=values.1dead, both_dead=values.2dead, ambiguous=values.ambiguous, bed=values.bed, true=values.true)

knitr::kable(data)
# intersect(gene.pair.types[!gene.pair.types$type %in% c("ambiguous"),], protein.pair.types[!protein.pair.types$type %in% c("ambiguous"),])
# #dead expression and still dead abundance (17/297)
# protein.pair.types[protein.pair.types$Maize1 %in% gene.pair.types$Maize1[gene.pair.types$type %in% c("1dead","2dead")],] %>% subset(type %in% c("1dead","2dead"))
```

@@

```{r print, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
# # ggsave("Output/Figure 1a.png", plot = chromosome_plot_exp)
# png(filename = "Output/Figure 1a.png", type = "cairo", units = "in", width = 8, height = 8, pointsize = 12, res = 96)
# plot(chromosome_plot_exp)
# dev.off()
# 
# # ggsave("Output/Figure 1b.png", plot = chromosome_plot_abundance)
# png(filename = "Output/Figure 1b.png", type = "cairo", units = "in", width = 8, height = 8, pointsize = 12, res = 96)
# plot(chromosome_plot_abundance)
# dev.off()
# 
# png("Output/Figure 2a.png");
# grid.draw(GOTerms_by_subgenome_a)
# dev.off()
# png("Output/Figure 2b.png");
# grid.draw(GOTerms_by_subgenome_b)
# dev.off()
# 
# # ggsave("Output/Figure 3.png", plot = subgenome_isoforms_scatter)
# png(filename = "Output/Figure 3.png", type = "cairo", units = "in", width = 8, height = 8, pointsize = 12, res = 96)
# plot(subgenome_isoforms_scatter)
# dev.off()
# 
# # ggsave("Output/Figure 4.png", plot = subgenome_expression_scatter_rep)
# png(filename = "Output/Figure 4.png", type = "cairo", units = "in", width = 12, height = 4, pointsize = 12, res = 96)
# plot(subgenome_expression_scatter_rep)
# dev.off()
# 
# # ggsave("Output/Figure 5.png", plot = subgenome_protein_scatter_rep)
# png(filename = "Output/Figure 5.png", type = "cairo", units = "in", width = 12, height = 4, pointsize = 12, res = 96)
# plot(subgenome_protein_scatter_rep)
# dev.off()
# 
# # ggsave("Output/Figure 6.png", plot = subgenome_expression_protein_scatter_rep)
# png(filename = "Output/Figure 6.png", type = "cairo", units = "in", width = 12, height = 4, pointsize = 12, res = 96)
# plot(subgenome_expression_protein_scatter_rep)
# dev.off()
# 
# ggsave("Output/Figure S1.png", plot = subgenome_expression_scatter)
# ggsave("Output/Figure S2.png", plot = subgenome_protein_scatter)
# ggsave("Output/Figure S3.png", plot = subgenome_expression_protein_scatter)
# ggsave("Output/Figure S4.png", plot = subgenome_expression_protein_foldchange_scatter)
# 
# list.of.subgenome.assignments <- subgenome %>% select(gene2, subgenome) %>% rename(B73.v4.ID=gene2, Subgenome=subgenome)
# write.csv(list.of.subgenome.assignments, "list.of.subgenome.assignments.csv")
# 
# write.csv(homeologs.pairs %>% subset(Maize1 != "" & Maize2 != "") %>% select(Maize1, Maize2) %>% rename(B73.v4.ID.Maize1=Maize1, B73.v4.ID.Maize2=Maize2), "list.of.retained.duplicate.pairs")
```
